\input{../meta.tex}
\newcommand{\soptitle}{Design and Analysis of Symmetric Primitives for Zero-knowledge Protocols}

\title{\sffamily\bfseries\Large\textcolor{theme}{\soptitle}}\par
\author{
    Shibam Ghosh\textsuperscript{1}\\[1ex]
    \normalsize \textsuperscript{1}Indian Institute of Technology Hyderabad 
}

\date{} % Leave empty to hide the date, or remove this line

\begin{document}
% ---------- Title Box ----------
\begin{tcolorbox}[
    colback=headerbg,
    colframe=headerbg,
    boxrule=0pt,
    width=\textwidth,
    arc=0mm,
    left=6pt,
    right=6pt,
    top=6pt,
    bottom=6pt
]
\centering
{\LARGE \textbf{SPERK: Symmetric Primitives with Efficient Zero-Knowledge Proofs}}\\[6pt]

{\large \textbf{Principal Investigator: Shibam Ghosh}}\\
{\large \textbf{Position: Assistant Professor}}\\ 
{\large \textbf{Joining Date: 27th May, 2026}}\\ 
{\large Department of Computer Science and Engineering}\\[4pt]

Email: \href{mailto:shibam@cse.iith.ac.in}{shibam@cse.iith.ac.in}
\;|\;
Web: \href{https://shibamcrs.github.io/}{https://shibamcrs.github.io/}
\end{tcolorbox}

\vspace{8pt}

\section{Introduction}
Zero-knowledge proofs (ZKPs), since their seminal introduction in 1985 by Goldwasser, Micali, and Rackoff~\cite{DBLP:conf/stoc/GoldwasserMR85}, have fundamentally transformed the landscape of cryptographic protocol design.
ZKPs enable a prover to convince a verifier of possessing a witness $w$ such that $F(x; w) = y$, without revealing $w$.
This remarkable property balances privacy and verifiability, enabling secure information exchange while keeping sensitive data confidential. Traditionally, cryptography has focused primarily on protecting data confidentiality and integrity.
By contrast, ZKPs shift the emphasis toward verifiable computation integrity: they allow one to prove that a function was evaluated correctly, without disclosing the underlying inputs.
\textbf{A key question, then, is what kinds of computations should be prioritized?}
This research proposal focuses on the efficient zero-knowledge proof of computations involving symmetric-key cryptographic primitives, such as hash functions, block ciphers, and pseudorandom functions (PRFs).

% ───────────────────────────────────────────────────────────────────────
\subsection{Use-Inspired Context}
% ───────────────────────────────────────────────────────────────────────
The far-reaching implications of ZKPs have led to their widespread adoption and investigation within both academia and industry.
Their application domains are diverse and rapidly evolving, encompassing blockchain privacy, solving the scalability issues in decentralized ledgers, verifiable storage, and cross-chain interoperability.
In the financial and privacy-focused cryptocurrency sector, ZKPs are central to projects such as Zcash~\cite{zcash} and Aztec~\cite{aztec}, which uses zk-SNARKs to enable completely private blockchain transactions; StarkWare~\cite{starkware} and zkSync~\cite{zksync} (by Matter Labs), which employ ZKPs to provide scalable ZKP-powered layer-2 solutions on Ethereum~\cite{ethereum}; and and Avail~\cite{avail}, which uses ZKPs to provide verifiable data availability guarantees and support cross-chain interactions. 
Beyond blockchain technology, ZKPs are increasingly recognized as critical components in systems such as electronic voting (Helios~\cite{helios}), privacy-preserving authentication (Microsoft Azure~\cite{azureprivacy}, Polygon ID~\cite{polygonid}, Google Wallet~\cite{googlewallet}), delegating computations to untrusted servers and even privacy-enhanced machine learning.

\subsection{Problem Statement}
Despite rapid advances, significant gaps remain between the theoretical potential of ZKPs and their practical deployment, especially where efficient proofs of computation involving symmetric primitives are needed, which are critically important due to their broad practical applications.
For example, any ZKP that relies on Merkle proofs necessitates building hash function circuits within the proof system.
This is central to rollup-protocol-based solution for blockchain scalability like zkSync Era~\cite{zksync} and StarkNet~\cite{starkware}, where ZKPs are employed to demonstrate the correctness of updates to the rollup's Merkleized state.
Another emerging application areas of ZKPs are \textbf{post-quantum secure digital signatures} based on MPC-in-the-Head (MPCitH)~\cite{DBLP:conf/uss/GiacomelliMO16} and VOLE-in-the-Head (VOLEitH)~\cite{DBLP:conf/crypto/BaumBGKORS22} frameworks.
In such a framework, given a one-way function $f$ (such as a block cipher) and its output $y$ over a secret $x$, i.e., $y = f(x)$, one can construct a digital signature scheme with secret key $x$, public key $y$, and provides a non-interactive ZKP of the knowledge (NIZKPoK) of the secret $x$ as a signature.
However, mainstream primitives like \tx{AES} and \tx{SHA-3} were not designed for such proof systems, often leading to inefficient proofs and high performance overheads due to a mismatch between their internal structures and ZKP requirements.
ZKPs prioritize simple circuit representations of these primitives with simple algebraic structures rather than traditional efficiency metrics for software or hardware implementations.
As a result, there is a growing demand for ``ZK-friendly'' symmetric-key primitives that maximize efficiency and compatibility within ZKP environments.

% \subsection{Challenges}
% At the heart of all modern ZKP systems lies a shared foundational principle: how to represent computations as arithmetic circuits over finite fields, a process known as \emph{arithmetization}.
% Arbitrary computations are represented by systems of low-degree multivariate polynomials, preserving both the correctness of computation and the privacy of inputs.
% Once computation has been recast in polynomial form, modern protocols can leverage well-studied interactive proof systems to check the consistency of these polynomials.
% In doing so, they establish not only that a particular computation has been performed correctly, but also that sensitive information remains unexposed.
% Powerful algebraic tools, such as polynomial commitment schemes and error-correcting codes, can then be employed to realize highly efficient proof generation and verification under this model.

% However, arithmetization typically imposes significant overhead in proof size, generation, and verification time.
% Therefore, ZKP systems favor circuit representations that minimize algebraic complexity.
% A prime example is Rank-1 Constraint System (R1CS)~\cite{DBLP:journals/cacm/ParnoHG016} which model computations as a set of quadratic constraints, each of the form $L(x) \cdot R(x) = O(x)$ where $x$ represents input and intermediate variables, and $L, R, O$ matrices.
% This approach breaks the computation into addition and multiplication gates over a prime field (to get the bilinear-pairing property of the underlying elliptic curve), with the total number of multiplication gates directly tied to the number of R1CS constraints.
% As these dominate proof size and generation time, minimizing multiplication (and multiplication depth) becomes critical for efficiency.
% Another key arithmetization model is the Algebraic Intermediate Representation (AIR), where a computation is captured as an execution trace: a sequence of internal program states linked by low-degree polynomial equations of degree $d$.
% AIR can succinctly represent both Turing machines and RAM machines, with each state comprising $\mx{w}$ field elements and undergoing $\mx{T}$ transformations.
% Proof generation complexity scales with the value of $\mx{w\cdot d \cdot T}$, making it essential to minimize state transitions.
% ─────────────────────────────────────────────────────────────────────────────
\subsection{Objectives}
% ─────────────────────────────────────────────────────────────────────────────
This research project aims to resolve the fundamental tension between security and efficiency in symmetric-key primitives designed for advanced cryptographic protocols.
The core objective is to systematically investigate, design, and formalize secure primitives tailored for Zero-Knowledge Proof (ZKP) systems and Multi-Party Computation (MPC), where algebraic simplicity often leads to vulnerabilities. This will be achieved through a multi-pronged approach:

\begin{enumerate}[label=\textbf{O\arabic*.}, leftmargin=*, itemsep=2pt, start=1]
    \item  \textbf{Cryptanalysis:} Rigorously evaluating existing ZKP-friendly designs (e.g., MiMC, Poseidon, Polocolo) using advanced algebraic methods, including Gr\"obner basis and novel FFT-based attacks, to expose their security weaknesses.
    \item \textbf{Design:} Creating and analyzing novel MPC-friendly tweakable block ciphers to significantly improve the efficiency and reduce the signature size of post-quantum signature schemes like MPCitH, while also investigating their resilience against physical attacks (fault injection, side-channels) in hardware.
    \item \textbf{Formalization:} Developing a comprehensive theoretical framework and robust design guidelines for creating ``ZK-friendly'' primitives, ensuring their security and compatibility across diverse ZKP systems like R1CS and AIRs
\end{enumerate}

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Broader Impact}
% ─────────────────────────────────────────────────────────────────────────────

Beyond the immediate academic contributions, the proposed research carries significant strategic and societal implications, particularly in establishing technological self-reliance and securing next-generation infrastructure against real-world threats.

\begin{enumerate}[label=\textbf{I\arabic*.}, leftmargin=*, itemsep=2pt, start=1]

    \item \textbf{Fostering Indigenous Cryptographic Capabilities for Advanced Protocols:}
As advanced cryptographic protocols like Zero-Knowledge Proofs (ZKPs) and Multi-Party Computation (MPC) become the backbone of next-generation digital public infrastructure, privacy-preserving machine learning, and decentralized networks, reliance on external or legacy cryptographic standards introduces strategic dependencies. By successfully designing and rigorously evaluating a novel, indigenous MPC-friendly tweakable block cipher, this project directly contributes to national technological sovereignty. 

    \item \textbf{Securing Post-Quantum Digital Signatures Against Physical Exploits:}
The global transition to Post-Quantum Cryptography (PQC) is accelerating, with MPC-in-the-Head (MPCitH) based digital signatures emerging as highly efficient and robust candidates. However, mathematical security guarantees are rendered void if the physical hardware executing these algorithms is compromised. As these new PQC signatures are deployed in physically accessible and resource-constrained environments, such as smart cards, IoT devices, automotive systems, and cryptocurrency hardware wallets, they will inevitably face sophisticated side-channel and fault injection attacks. By providing a proper, systematic analysis against physical attacks, this research bridges a critical gap in current PQC deployment.
\end{enumerate}


\begin{genericbox}{Summary}[teal!60]
This project is dedicated to the design and analysis of symmetric-key tailored for efficient integration with zero-knowledge proof systems.
Addressing this specialized requirements presents both opportunities and challenges for cryptography research.
My expertise in design and cryptanalysis of symmetric-key primitives is well suited to investigating these challenges, systematically balancing security with application-specific performance.
Through this research, I aim to advance both the theoretical foundations and practical deployment of tailored primitives, bridging the gap between robust security guarantees and efficiency.
The resulting analysis and frameworks are intended to directly support the work of the standardization community, providing the rigorous security analysis and clear metrics needed for its technical examinations.
\end{genericbox}

% ─────────────────────────────────────────────────────────────────────────────
\section{Research Approach \& Feasibility}
% ─────────────────────────────────────────────────────────────────────────────

The research is structured into three temporally ordered thrusts, progressing from problems where established techniques provide a clear starting point (short-term) to problems that demand fundamentally new conceptual frameworks (long-term).
Feasibility at each stage is grounded on my existing expertise and preliminary results.

\subsection{Thrust~1 (Short-term): Cryptanalysis of Existing ZK-friendly Primitives}
Recent years have seen a surge in the proposal of new primitives tailored for zero-knowledge proofs, such as \tx~{MiMC}~\cite{DBLP:conf/asiacrypt/Albrecht0R0T16}, the \tx{Marvellous} family~\cite{DBLP:journals/tosc/AlyABDS20}, and \tx{Poseidon}~\cite{DBLP:conf/uss/0001KR0S21} to name a few.
The trend towards algebraic simplicity in those design significantly alters the cryptanalytic landscape.
Classical statistical attacks, such as differential and linear cryptanalysis, become less effective due to the flexibility of field sizes and rapid mixing with few rounds.
Instead, as the algebraic representation of these ciphers becomes more concise, their susceptibility to algebraic attacks such as interpolation, GCD attacks, and, most critically, Gröbner basis methods grows.
Notably, features that aid efficient proof generation (concise polynomial systems) simultaneously assist algebraic attack algorithms, creating a fundamental tension in balancing efficiency and security.
This has led to many proposed constructions later failing to achieve their claimed security, highlighting the need for more extensive cryptanalysis in this area.
Currently, I am investigating the Gr\"obner basis attacks on the \tx{Polocolo} hash function~\cite{DBLP:journals/iacr/HaHLPS25}, designed for R1CS and PLONK-based proof systems.
I am also exploring the possibility of applying FFT-based key recovery attacks that I have proposed in~\cite{DBLP:conf/eurocrypt/DunkelmanGKLMM24} for \tx{AES}, on the ciphers defined over prime field ($\mathbb{F}_p^n$) such as \tx{AES-prime}~\cite{DBLP:conf/eurocrypt/MasureMMS23}, which was proposed for efficient use in MPC protocols.

\subsection{Thrust~2 (Medium-term): Design and Analysis of MPC-friendly Primitives}
A key distinction of MPCitH post-quantum signature schemes is that their security relies solely on the underlying one-way function, providing an alternative hardness assumption for post-quantum cryptography.
This approach avoids the reduction gaps found in schemes based on specific mathematical problems, making them generally more robust.
While these signatures do not produce succinct proofs like SNARKs, they are highly efficient for small circuits, such as those used in digital signatures.
Notably, 6 out of 14 second-round candidates in the 2023 NIST post-quantum competition employ such constructions.
These schemes mainly differ in their choice of underlying one-way function.
Options range from symmetric primitive-based functions to problems like syndrome decoding, subset sum, and multivariate quadratic equations.
While most are built on well-studied security assumptions, their general performance lags behind symmetric primitive-based approaches, which thus attract significant interest.
A challenge, however, is that standard symmetric primitives are not as efficient for this setting as the custom one-way functions designed for MPCitH.
For example, the well-known \tx{Picnic} signature~\cite{DBLP:conf/ccs/ChaseDGORRSZ17} uses the MPC-friendly block cipher \tx{LowMC}~\cite{DBLP:conf/eurocrypt/AlbrechtR0TZ15}, which minimizes the number of \txf{AND} gates.
Still, optimizing for efficiency presents new security challenges, as \tx{LowMC} has faced several devastating attacks.
This has driven recent schemes, such as \tx{BBQ}~\cite{DBLP:conf/sacrypt/GuilhemMOS19}, \tx{Banquet}~\cite{DBLP:conf/pkc/BaumGKOSZ21}, and \tx{FAEST}~\cite{DBLP:conf/crypto/BaumBGKORS22}, to be based on standard primitives like \tx{AES}.
However, in this context, attackers typically have access to only one plaintext-ciphertext pair, suggesting that tailored primitives could offer significant gains in signature size and performance, a key motivation for this project.
Within this research, I will pursue two main directions:

\subsubsection{Designing MPC-friendly Tweakable block ciphers.} 
With block ciphers as the underlying one-way function, current approaches require changing the master key for every signature, making an MPC-friendly key schedule essential and increasing overhead in both efficiency and signature size. 
Instead one can benefit from using a tweakable block cipher by changing only the tweak per signature.
Since tweaks are public, their scheduling need not be MPC-friendly, enabling the adoption of more complex, non-linear algorithms.
Our previous work on nested tweakey schedule for \tx{ChiChi}~\owncite{DBLP:conf/eurocrypt/BelkheyarDGLMPRSSAV25} distinctly separate key and tweak scheduling, and I plan to extend this strategy to design new tweakable block ciphers optimized for such applications.
This MPC-friendly design also enhances resilience against side-channel attacks through masking over groups of prime order.
While Boolean masking is a common countermeasure, its effectiveness is often limited by practical challenges in maintaining sufficiently noisy leakages (especially for small embedded devices).
Dziembowski et al.~\cite{DBLP:conf/tcc/DziembowskiFS16} demonstrated that masking over groups of prime order group can benefit from the mismatch between real-world leakage functions and the algebraic structure of the masking field.
As a result, this field is now rapidly emerging, and we have recently seen several new constructions proposed~\cite{DBLP:conf/eurocrypt/MasureMMS23, DBLP:conf/eurocrypt/GrassiMMMS24}.

\subsubsection{Physical Attacks.} 
Ensuring resistance to physical attacks such as fault injection and side-channel leakage is critical, especially as these signature schemes are often implemented in hardware (e.g., cryptocurrency cold wallets).
Recent work has revealed vulnerabilities in components used like polynomial commitment scheme~\cite{DBLP:conf/pqcrypto/GodardAGLM25,DBLP:journals/iacr/JendralD25} and one-way function~\cite{cryptoeprint:2025/1895}.
However, this field of physical attacks has yet to be explored more deeply, especially for protocols utilizing MPC-friendly primitives, such as \txf{AIMer}~\cite{DBLP:conf/ccs/KimHSLMLLKCY023}, the winner of the Korean Post-Quantum Cryptography Competition.
This line of research will identify the most vulnerable components within these protocols and guide the development of effective protection strategies using established countermeasures, such as masking or duplication.

\subsection{Thrust~3 (Long-term): Fomalization of Design Criteria for ZK-friendly Primitives}
The practical realization of ZKPs advanced significantly with the advent of zero-knowledge succinct non-interactive arguments of knowledge (zk-SNARKs), first introduced by Ben-Sasson et al.~\cite{DBLP:conf/uss/Ben-SassonCTV14} in 2013.
Since then, the field has rapidly expanded to incorporate diverse zk-SNARK protocols --- such as \txf{Pinocchio}~\cite{DBLP:journals/cacm/ParnoHG016}, \txf{Groth16}~\cite{DBLP:conf/eurocrypt/Groth16}, \txf{PLONK}~\cite{DBLP:journals/iacr/GabizonWC19}, and many others, as well as Zero-Knowledge Scalable Transparent Arguments of Knowledge (zk-STARKs)~\cite{DBLP:conf/crypto/Ben-SassonBHR19}.
Notably, different ZKP frameworks use distinct algebraic models depending on their proof system and efficiency goals.
For example, zk-SNARKs typically employ Rank-1 Constraint Systems (R1CS) with Quadratic Arithmetic Programs (QAPs), while zk-STARKs utilize Algebraic Intermediate Representations (AIRs).
Each framework presents unique trade-offs in proof size, verification speed, trust assumptions, and compatibility with various types of symmetric-key primitives.
My long-term objective is to develop a comprehensive formalization of the fundamental principles underpinning these frameworks and to articulate robust design guidelines for arithmetization-friendly primitives.
By deepening the theoretical foundations of algebraic cryptanalysis, including refining existing methods such as Gr\"obner basis attacks, resultant computations, I aim to enable both rigorous security evaluation and efficient design.

% ─────────────────────────────────────────────────────────────────────────────
\section{Expected Deliverables}
% ─────────────────────────────────────────────────────────────────────────────
\subsection*{Thrust 1 — Cryptanalysis of Existing Ciphers}
\begin{enumerate}[label=\textbf{D\arabic*.}, leftmargin=*, itemsep=2pt]
  \item \textbf{Cryptanalytic results on deployed ciphers.}
        Published attacks in terms of finding new distinguishers and key-recovery results on at least two of the target ciphers, targeting publication at CRYPTO, EUROCRYPT, ASIACRYPT, or IACR ToSC.

  \item \textbf{High performance cryptanalysis using GPU.}
      Cryptanalysis of ZK-friendly block ciphers often required solving system of non-linear equations over large finite fields.
      We will develop GPU-accelerated solvers for these equations, enabling fast and efficient cryptanalysis of ZK-friendly ciphers.
        An open-source library will be made available for reproducible and extensible cryptanalysis by the wider community.
\end{enumerate}
\subsection*{Thrust 2 — Design and Analysis of MPC-friendly Primitives}
\begin{enumerate}[label=\textbf{D\arabic*.}, leftmargin=*, itemsep=2pt]
  \setcounter{enumi}{2} % Continues numbering from D2
  \item \textbf{Novel MPC-friendly Tweakable Block Cipher.} 
        The design, specification, and comprehensive security evaluation of a new tweakable block cipher optimized for MPC-in-the-Head (MPCitH) signature schemes and demonstrating concrete improvements in signature size and verification efficiency compared to AES-based or LowMC-based alternatives.
        
  \item \textbf{Vulnerability Reports and Countermeasure Guidelines for Physical Attacks.}
        Detailed physical security evaluations (side-channel and fault injection) of hardware implementations for prominent MPCitH schemes (e.g., AIMer). Deliverables will include public vulnerability reports and a framework of effective protection strategies.
\end{enumerate}

\subsection*{Thrust 3 — Formalization of Design Criteria}
\begin{enumerate}[label=\textbf{D\arabic*.}, leftmargin=*, itemsep=2pt]
  \setcounter{enumi}{4} % Continues numbering from D4
  \item \textbf{Robust Design Guidelines for Arithmetization-Friendly Primitives.} 
        A comprehensive formalization detailing the design criteria required to balance efficiency and security across different ZKP frameworks (e.g., R1CS for zk-SNARKs and AIRs for zk-STARKs). This will serve as a foundational manual for future cryptographic designers.
        
  \item \textbf{Theoretical Advancements in Algebraic Cryptanalysis.}
        Publication of refined theoretical methods for algebraic cryptanalysis, specifically improving the efficiency and applicability of Gr\"obner basis and resultant-based attacks on algebraically simple primitives. 
\end{enumerate}

% ─────────────────────────────────────────────────────────────────────────────
\section{Budget Allocation}
% ─────────────────────────────────────────────────────────────────────────────
The proposed budget of \textbf{Rs. 34.44 Lakhs} over two years is carefully planned to meet the project’s specific computational, hardware, manpower, and dissemination requirements. Here is the detailed budget breakdown:


\begin{table}[htbp]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{} l p{9.5cm} r @{}}
\hline
\textbf{Category} & \textbf{Description} & \textbf{Amount (Rs.)} \\
\hline
\textbf{Equipment}
& \textbf{High-Performance Workstation:} \newline High-end Desktop Workstation (e.g., Dell Precision/HP Z series), top-tier multi-core processor (Intel Xeon/Core i9 or AMD Threadripper), 128GB RAM, 2TB NVMe SSD + 4TB HDD, and a high-compute GPU (e.g., NVIDIA RTX 4090 / A6000 class). & 3,50,000 \\
& \textbf{Hardware Security Setup:} \newline Fault injection and side-channel analysis tools (e.g., ChipWhisperer hardware, Electromagnetic Fault Injection (EMFI) tools, high-speed oscilloscope, and FPGA evaluation boards). & 7,82,000 \\
\cline{2-3}
& \textit{Equipment Subtotal} & \textit{11,32,000} \\
\hline
\textbf{Manpower} & \textbf{Ph.D. Scholar Fellowship:} \newline Support for one Ph.D. student for up to four years (Rs. 36,000/month avg.). & 17,28,000 \\
\hline
\textbf{Travel} & Conference presentations, networking, and collaborative visits. & 4,00,000 \\
\hline
\textbf{Contingency} & Target microcontrollers, probing consumables, software licenses, maintenance, and open-access publication charges. & 2,00,000 \\
\hline
\textbf{Total} & & \textbf{34,60,000} \\
\hline
\end{tabular}
\caption{Detailed Breakdown of the Proposed Budget}
\label{tab:budget}
\end{table}

\begin{enumerate}[label=\textbf{B\arabic*.}, leftmargin=*, itemsep=4pt]
  \item \textbf{Human Resource Development (Rs. 17.28 Lakhs):}
        This allocation supports the fellowship of one Ph.D. student for four years. The researcher will be trained in the highly specialized intersection of symmetric-key cryptography, automated cryptanalysis, and hardware security, directly contributing to the project deliverables.

  \item \textbf{Computational \& Hardware Resources (Rs. 11.32 Lakhs):}
        The cryptanalysis of ZK-friendly ciphers often requires solving massive systems of non-linear equations (Gr\"obner basis) and running resource-heavy automated solvers, necessitating a high-end, GPU-accelerated desktop workstation. Furthermore, to evaluate physical security (fault-injection and side-channel attacks), a dedicated hardware testing setup is required. This includes tools like high-speed oscilloscopes, Electromagnetic Fault Injection (EMFI) equipment, and FPGA boards to accurately mimic and attack real-world hardware deployments.

  \item \textbf{Travel and Dissemination (Rs. 4.00 Lakhs):}
        Allocated to promote outreach and collaboration. This will support the Principal Investigator and the Ph.D. student in presenting their findings at top-tier international cryptographic conferences (e.g., CRYPTO, EUROCRYPT, CHES, FSE) and facilitate collaboration visits with global experts in MPC and ZKP primitives.

  \item \textbf{Contingency and Consumables (Rs. 2.00 Lakhs):}
        Reserved for day-to-day experimental consumables (such as target embedded boards, cables, and probing needles for physical attacks), software licenses, system maintenance, and open-access publication fees to ensure the research is freely accessible to the global community.
\end{enumerate}

\begin{figure}[htbp]
    \centering
    % Define custom colors to match the image exactly
    \definecolor{barorange}{RGB}{237, 125, 49}
    \definecolor{barpink}{RGB}{233, 163, 227}
    \definecolor{barblue}{RGB}{0, 176, 240}
    \definecolor{bargreen}{RGB}{0, 176, 80}
    \definecolor{textgray}{RGB}{90, 90, 90}
    \definecolor{gridgray}{RGB}{220, 220, 220}

    % Adjust x and y scaling here to fit your document
    \begin{tikzpicture}[scale=0.8, x=0.6cm, y=0.7cm, font=\sffamily\footnotesize]

        % 1. Draw Vertical Grid Lines
        \draw[gridgray, thin] (0, 0.5) -- (0, 10.5);
        \foreach \i in {1,...,24} {
            \draw[gridgray, thin] (\i, 0.5) -- (\i, 10.5);
        }

        % 2. Draw X-axis Numbers (Months)
        \foreach \i in {1,...,24} {
            \node[anchor=north, text=textgray, font=\sffamily\scriptsize\bfseries] at (\i, 0.5) {\i};
        }

        % 3. Draw the Outer Border
        \draw[gridgray, thick] (-9, 0.5) rectangle (24.5, 10.5);

        % 4. Define a macro for drawing tasks to keep the code clean
        % Usage: \task{y-position}{Start Month}{End Month}{Color}{Task Name}
        % Note: Month 1 starts at x=0 and ends at x=1.
        \newcommand{\task}[5]{
            % Task Label
            \node[anchor=east, text=textgray, font=\sffamily\scriptsize\bfseries] at (-0.2, #1) {#5};
            % Task Bar
            \fill[#4] (#2, #1 - 0.2) rectangle (#3, #1 + 0.2);
        }

        % 5. Add the Tasks (from bottom to top, matching your image style)
        % y=1 is the bottom task, y=9 is the top task

        \task{1}{0}{3}{barorange}{Equipment Procurement}
        \task{2}{0}{4}{barorange}{Human Resource Hiring}

        \task{3}{3}{8}{barpink}{Cryptanalysis Project}

        \task{4}{6}{12}{barblue}{GPU-accelerated Solver}

        \task{5}{11}{12}{bargreen}{Mid Term Project Report}

        \task{6}{10}{17}{barblue}{Design of MPC-friendly Ciphers}

        \task{7}{14}{21}{barblue}{Physical Attack Analysis}

        \task{8}{17}{23}{barblue}{Formalization of Design Criteria}

        \task{9}{23}{24}{bargreen}{Final Project Report}

    \end{tikzpicture}

    \caption{24 Months Timeline}
    \label{fig:timeline}
\end{figure}

% ─────────────────────────────────────────────────────────────────────────────
\section{Conclusion}
% ─────────────────────────────────────────────────────────────────────────────
This project addresses a critical bottleneck in modern cryptographic engineering: the fundamental tension between the extreme efficiency required by advanced protocols, such as Zero-Knowledge Proofs and Multi-Party Computation, and their resilience against mathematical and physical attacks.
The Center for Cryptography and Cybersecurity at the Indian Institute of Technology (IIT) Hyderabad offers an outstanding environment for advancing research in these fields.
My research plan is closely aligned with the center’s core areas, including symmetric key cryptography and cryptanalysis, post-quantum cryptography (PQC), and zero-knowledge proofs in blockchain environments.

\pagebreak
\bibliographystyle{alpha}
\bibliography{../bibliography.bib,../myPublications.bib}
\end{document}
